---
title: "RWorksheet_Porras#4C"
author: "Porras, Christian BSIT 2-C"
date: "2023-11-23"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1.A Solutions on how to import a csv file into the environment.
```{r}

mpgdata <- read.csv("mpg.csv")
mpgdata

```


1.B variables from mpg dataset are categorical?
```{r}
str(mpgdata)
str

```


1.C Which are continuous variables?
```{r}
#displ: Engine displacement in liters.
#hwy: Highway miles per gallon.
#cty: City miles per gallon.
```

2. Which manufacturer has the most models in this data set? Which model has the most variations? Show your answer
```{r}
library(ggplot2)


mostmod <- table(mpg$manufacturer)

maxmod <- names(which.max(mostmod))

cat("Manufacturer with the most models:", maxmod, "\n")


#finding most variations
maxdata <- subset(mpg, manufacturer==maxmod)
modcounts <- table(maxdata$model)

 variation_model <- names(which.max(modcounts))
cat("Model with the most variations:",variation_model, "\n")

```

A.Group the manufacturers and find the unique models. Show your codes and result.
```{r}
library(ggplot2)
library(dplyr)

unimode <- mpg%>%
  group_by(manufacturer)%>%
  summarize(uniquemodel = n_distinct(model))

print(unimode)

```

B. Graph the result by using plot() and ggplot(). Write the codes and its result.
```{r}

#ggplot
library(ggplot2)
ggplot(unimode, aes(x=manufacturer,y = uniquemodel))+
  geom_bar(stat = "Identity",fill = "blue")+
  labs(title = "Unique model per manufacturer",
       x = "Manufacturer",
       y = "Number of unique models")+
  theme_gray()

```

2. Same dataset will be used. You are going to show the relationship of the modeland the manufacturer

A.What does ggplot(mpg, aes(model, manufacturer)) + geom_point() show?
```{r}
library(ggplot2)
ggplot(mpg, aes(model, manufacturer)) + geom_point()

# gplot(mpg, aes(model, manufacturer)) + geom_point() shows the points where model and manufacturer.

```

B. For you, is it useful? If not, how could you modify the data to make it more informative?
```{r}
#yes it is

```

3. Plot the model and the year using ggplot(). Use only the top 20 observations. Write the codes and
its results.
```{r}
library(ggplot2)

top20 <- head(mpg,20)

ggplot(top20, aes(x = model, y = year))+
  geom_point()+
  labs(title = "Model vs Year(Top 20 observations)",
       x = "Model",
       y= "Year")+
  theme(axis.text.x = element_text(angle = 45,hjust = 1))
```



4.Using the pipe (%>%), group the model and get the number of cars per model. Show codes and its result
```{r}
library(dplyr)
group_model <- mpg%>%
  group_by(manufacturer)

print(group_model)

```

A. Plot using geom_bar() using the top 20 observations only. The graphs shoudl have a title, labels and
colors. Show code and results.
```{r}
library(dplyr)
library(ggplot2)

data(mpgdata)

top_observe <- head(mpg,20)

ggplot(top_observe, aes(x = model, fill = model)) +
  geom_bar() +
  labs(title = "Number of Cars per Model (Top 20 Observations)",
       x = "Model",
       y = "Number of Cars") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "none") +
  scale_fill_manual(values = rainbow(nrow(top_observe))) 


```


B. Plot using the geom_bar() + coord_flip() just like what is shown below. Show codes and its result.
```{r}
library(dplyr)
library(ggplot2)

data(mpgdata)

top_observe <- head(mpg,20)

ggplot(top_observe, aes(x = model, fill = model)) +
  geom_bar() +
  labs(title = "Number of Cars per Model (Top 20 Observations)",
       x = "Number of Cars",
       y = "Model") +
  theme_minimal() +
  theme(legend.position = "none") +
  scale_fill_manual(values = rainbow(nrow(top_observe))) +
  coord_flip()



```

5. Plot the relationship between cyl - number of cylinders and displ - engine displacement using
geom_point with aesthetic color = engine displacement. Title should be “Relationship between
No. of Cylinders and Engine Displacement”.
```{r}
library(ggplot2)

data(mpgdata)


ggplot(mpg, aes(x = cyl, y = displ, color = displ)) +
  geom_point() +
  labs(title = "Relationship between No. of Cylinders and Engine Displacement",
       x = "Number of Cylinders",
       y = "Engine Displacement") +
  scale_color_continuous(name = "Engine Displacement") +
  scale_color_gradient(low="blue",high="red") +
  theme_minimal()

```

A. How would you describe its relationship? Show the codes and its result.
```{r}
relation_cyldispl <- cor(mpg$cyl , mpg$displ)

cat("Correlation Coefficient:", relation_cyldispl, "\n")


library(ggplot2)
ggplot(mpg, aes(x = cyl, y = displ, color = displ)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, color = "red") +
  labs(title = "Relationship between No. of Cylinders and Engine Displacement",
       x = "Number of Cylinders",
       y = "Engine Displacement") +
  scale_color_continuous(name = "Engine Displacement") +
  theme_minimal()

```

B. Plot the relationship between displ (engine displacement) and hwy(highway miles per gallon). Mapped
it with a continuous variable you have identified in #1-c. What is its result? Why it produced such
output?
 
```{r}

ggplot(mpg, aes(x = displ, y = hwy, color = year)) +
  geom_point() +
  labs(title = "Relationship between Engine Displacement and Highway MPG",
       x = "Engine Displacement",
       y = "Highway MPG") +
  scale_color_gradient(low="blue",high="red") +
  theme_minimal()


```


6. Import the traffic.csv onto your R environment.

A. How many numbers of observation does it have? What are the variables of the traffic dataset the Show
your answer.

```{r}
trafficdata <- read.csv("traffic.csv")

observation_count <- nrow(trafficdata)
var_traffic <- ncol(trafficdata)
names <- names(trafficdata)

cat("Number of Observations:", observation_count, "\n")
cat("Number of Variables:", var_traffic, "\n")
cat("Variables:", paste(names, collapse = ", "), "\n")

```


B. subset the traffic data set into junctions. What is the R codes and its output?
```{r}
junction <- c(1,2,3,4)
junction <- trafficdata[trafficdata$Junction %in% junction, ]
junction


```

C. Plot each junction in a using geom_line(). Show your solution and output.
```{r}
library(ggplot2)

ggplot(trafficdata, aes(x = DateTime, y = Vehicles, color = as.factor(Junction))) +
  geom_line() +
  scale_color_manual(values = c("red", "orange", "green", "blue")) + 
  labs(title = "Traffic by Junction",
       x = "Date Time",
       y = "Number of Vehicles") +
  theme_minimal() +
  facet_wrap(~Junction, scales = "free_y", ncol = 1) +
  guides(color = guide_legend(title = "Junctions"))
```

7. From alexa_file.xlsx, import it to your environment
```{r}
library(readxl)

alexafile <- read_excel("alexa_file.xlsx")
alexafile
```

A. How many observations does alexa file has? What about the number of columns? Show your solution
and answer.

```{r}
observation_num <- nrow(alexafile)

column_number <- ncol(alexafile)


cat("The number of observations on alexa is:",observation_num,"\n")
cat("The number of coloumns on alexa is:",column_number,"\n")

```


B.group the variations and get the total of each variations. Use dplyr package. Show solution and
answer.
```{r}
library(dplyr)
groupvar <- alexafile%>%
  group_by(variation)%>%
  summarise(totalcount_ = n())

groupvar

```
C. Plot the variations using the ggplot() function. What did you observe? Complete the details of the
graph. Show solution and answer.
```{r}
library(ggplot2)

ggplot(alexafile, aes(x = variation)) +
  geom_bar() +
  labs(title = "Distribution of Variations",
       x = "Variation",
       y = "Count") +
  theme_minimal()
```

D. Plot a geom_line() with the date and the number of verified reviews. Complete the details of the
graphs. Show your answer and solution.
```{r}
alexafile$date <- as.Date(alexafile$date)

alexafile$month <- format(alexafile$date, "%m")

monthscount <- alexafile %>%
  group_by(month) %>%
  summarise(num_reviews = n())


monthlyreview <- table(monthscount)

ggplot(monthscount, aes(x = month, y = num_reviews, group = 1)) +
  geom_line(color = "red") +
  labs(title = "Number of Verified Reviews Per Month",
       x = "Month of 2017", y = "Number of Reviews")

```

E. Get the relationship of variations and ratings. Which variations got the most h ghest in rating? Plot a
graph to show its relationship. Show your solution and answer.
```{r}
library(dplyr)
ggplot(alexafile, aes(x = variation, y = rating, fill = variation)) +
  geom_bar(stat = "summary", fun = "mean", position = "dodge") +
  scale_fill_manual(values = rainbow(n = length(unique(alexafile$variation)))) +  
  labs(title = "Relationship Between Variations and Ratings",
       x = "Variations",
       y = "Average Rating") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

var_ratings <-  alexafile %>%
  group_by(variation)%>%
  summarise(avearage_rating = mean(rating, na.rm = TRUE))

max_rating <- max(var_ratings$average_rating, na.rm = TRUE)

highrate <- alexafile %>%
  filter(rating == max_rating)
print(highrate)

```